---
title: "Video Laryngoscopy does not reduce intubation time in severly obese patients"
author: "Alexander Koch"
date:  "`r date()`"
format:
  html:
    toc: true
    code-fold: show
    code-tools: true
    df-print: paged
    embed-resources: true
  pdf:
    toc: true
---

```{r}
#| output: false
#| warning: false
# Load libraries

library(tidyverse)
library(lubridate)
library(ggplot2)
library(dplyr)
```

## Introduction

Difficult airways and restricted vision are a major challenge when intubating, even for experienced laryngoscopists. Usually, these problems are even aggravated in overweight patients. Novel video laryngoscopes are therefore increasingly being used in hospitals to facilitate quick and easy intubation in such situations. But do they deliver what they promise? To test this, Abdallah et. al \[1\] had 105 obese patients with a body mass index (BMI) between 30 and 50 kg/m^2^ intubated either with a standard Macintosh laryngoscope or a video laryngoscope by two experienced anesthesiologists. One major outcome was the total time required for successfull intubation.

## Study results

```{r}
#| warning: false
data <- medicaldata::laryngoscope

data_sel <- data %>% select(c("BMI", "Randomization", "total_intubation_time")) %>%
  mutate(Randomization = factor(Randomization))
```

Contrary to expectations, @fig-technique shows that total intubation time was higher when using a video laryngoscope.

```{r}
#| warning: false
#| label: fig-technique
#| fig-cap: "Total intubation time (sec) for control group (standard Macintosh laryngoscope) and intervention group (video-assisted laryngoscope)."
#| fig-width: 8
#| fig-height: 4
ggplot(data = data_sel, mapping = aes(x = Randomization, y = total_intubation_time, fill = Randomization)) + 
  geom_col() +
  labs(x = "", y = "Total intubation time (sec)") +
  scale_fill_manual(name = "Intubation technique",
                     values = c("skyblue3", "sienna2"),
                     labels = c("Standard", "Video")) +
  scale_x_discrete(labels=c('Standard', 'Video'))
```

@fig-BMI further shows that intubation even increasingly took longer the higher the BMI of the patient was. Surprisingly, this was not the case if using the standard laryngoscope.

```{r}
#| warning: false
#| label: fig-BMI
#| fig-cap: "Correlation of total intubation time (sec) and body-mass-index (BMI) for both control (standard Macintosh laryngoscope) and intervention group (video-assisted laryngoscope)."
#| fig-width: 8
#| fig-height: 4
ggplot(data = data_sel, mapping = aes(x = BMI, y = total_intubation_time, colour = Randomization)) +
  geom_point() +
  labs(x = "Body-mass-index (BMI)", y = "Total intubation time (sec)") +
  scale_color_manual(name = "Intubation technique",
                     values = c("skyblue3", "sienna2"),
                     labels = c("Standard", "Video")) +
  geom_smooth(method = "lm")
```

## Conclusions

The results clearly show that the time required for tracheal intubation using video laryngoscope was even longer than for the Macintosh laryngoscope. A video laryngoscope should therefore not routinely be substituted for a conventional one in morbidly obese patients. However, it should be considered that proper usage of a novel intubation technique always needs a certain amount of training and anesthesiologists might become more aquainted with video-assisted intubation in the future.

## References

\[1\] Abdallah R, Galway U, You J, Kurz A, Sessler DI, Doyle DJ. A randomized comparison between the Pentax AWS video laryngoscope and the Macintosh laryngoscope in morbidly obese patients. Anesth Analg. 2011 Nov;113(5):1082-7.
